Percussive Sound Generation with Virtual Listeners and Modular Synthesizers

### 2020-08-20 templates_aimusic2020/paper.tex
Updates from Overleaf

We opted for a set of classical DSP methods to build our synthesizer, which allow for fast offline generation of audio signals. We made extensive use of Pippi\footnote{https://github.com/luvsound/pippi} and SciPy~\cite{jones2001scipy} libraries. Our virtual synthesizer contains a set of one or more submodules. Each submodule is a self-contained noise making unit. Submodules have identical sets of parameters, but widely different outputs can be achieved depending on the values assigned. The set of parameters available to each submodule is highlighted in Table~\ref{table:submodule_params}. The sonic output of the virtual synthesizer is the normalized addition of the sonic output of its submodules. Our implementation of a synthesizer can have any number of submodules. The parameters that dictate the output signal of each submodule as well as the range of values each parameter can take are shown in table~\ref{table:submodule_params}. We call the number of submodules in each virtual synthesizer the \textit{stack size}. We call the sets of parameter values that characterize a synthesizer's submodules a \textit{program} (analogous to a preset for a VST).

@@

We are not looking for the perfect imitation of organic drums using a synthesizer. We seek to imitate drums using a synthesizer and even \textbf{prefer} for its generations to retain novel, unusual characteristics. The task assigned to the virtual ear is the rare acceptance of drum-like sounds and the inevitable rejection of most \enquote{noise} outputs from the virtual synthesizer. By definition, we cannot anticipate what these novel drums will sound like. Without manual inspection, all we know is that the accepted sounds are synthesizer noises which the synthetic ear has determined to be similar to drum-sounds. The virtual ear makes two critical decisions: 
\emph{
\begin{quote}
\text{Decision.1 Could the sound be used as a drum?}\label{Decision.1}
\\
\text{Decision.2 If it does sound like a drum, what type of drum should it be?}\label{Decision.2}
\end{quote}
 }

\begin{figure}[htbp]
    \begin{center}
    \textbf{Data Overview}
    \makebox[\textwidth]{
    \includegraphics[width=0.95\linewidth]{images/venn_data.pdf}}
    \end{center}
    \caption{ 
    Scales are symbolic (true scales are not measurable and/or subjective).  $\mathcal{N}$ is the set of percussive sounds a synthesizer is capable of making. Our positive samples, $\mathcal{T^{+}}$, is a small fraction of a wide variety of percussive sounds that are conceivable. The size of $\mathcal{T^{-}}$ can be set to any number(we can generate as any number of random samples), but we prefer to keep it no larger than  $\mathcal{T^{+}}$. $\mathcal{H}$ is a series of sounds sent to the ear for classification. The ideal ear will detect  $\mathcal{H^{+}}$, the overlap of $\mathcal{H}$ and $\mathcal{N}$, with high precision and recall. 
    }
\label{fig:ven_data}
\end{figure}
\hypref{Decision.1} requires knowledge of what drums \textbf{do not} sound like, or knowledge of an infinitely large set, which cannot be fully represented via examples. An important consideration is that the source of sounds used in training the model (organic drum sounds) will be fundamentally different from the source of unlabeled sounds we wish to categorize (noise from a synthesizer). This issue is reflective of the OSR problem~\cite{geng2020recent,mundt2019open}. Figure~\ref{fig:ven_data} highlights a number of caveats with this approach. we denote $\mathcal{N}$ as the set of percussive sounds a synthesizer is capable of making. $\mathcal{T^{+}}$, our positive examples, is a set of sound extracted from material drums. $\mathcal{T^{-}}$, is a small subset of an infinite set which is meant to represent all noises which the synthesizer is capable of making and likely includes sounds in $\mathcal{N}$. During the hearing test, the virtual ear continually receives synthesizer noises for classification. The set of received noises is denoted as $\mathcal{H}$. The performance of the ear is measured by its precision and recall in finding the overlap between $\mathcal{H}$ and $\mathcal{N}$. It is important to ensure that the change in learning domains ---particularly with the case of positive examples of percussive sounds--- does not interfere with transformation of knowledge from the training of the ear to the hearing test. To maximize the transference of knowledge from one domain to another, we seek to learn from agnostic feature sets that capture fundamental characteristics of the data points.
\subsection{Feature Extraction}
To train our virtual ear, we require generalizable, domain agnostic features. Various works have demonstrated effective reconstruction of signals given their Short-time Fourier Transforms (STFT) ~\cite{nawab1983signal,griffin1984signal}. If the STFT of a signal can be used its reconstruction, perhaps it can be utilized as a source of fundamental features necessary for audible signal categorization. We defined 3 STFT transformation functions which we believe to capture important, unique attributes of percussive sounds (figure ~\ref{fig:stackspectrums}). Our transformations are GPU based, allowing their application in tandem to the training of the classifier or tuning of STFT transformation during hyper-paramter optimization (see Appendix~\ref{appendix:hyperparam}).
\subsection{Ear Types}
\label{classifiers}

@@

\subsection{Survey Discussion}
\label{survey2_takeaway}
Due to the discrepancy in categorization and output groups, results from the two surveys must be compared with caution. Agree-ability scores for MEM pipeline are often lower, but we reiterate that the MEM survey has two out of six (rather than one out of six in the TPE pipeline) categories which are exclusive to survey responders. Supporting this hypothesis is the notable improvement in agree-ability observed after dropping all \enquote{bad} and \enquote{other} samples. Despite the drawbacks, the results here are encouraging as the MEM pipeline remains competitive despite working with a fraction of the features to learn from and simultaneously making both decisions. Suggesting that latent representations of autoencoder networks can be used as general, low-dimensional proxies for representation of complex inputs.

\bibliographystyle{apacite}
\bibliography{CSMC_MUME_LaTeX_Template.bib}

>>>>

We opted for a set of classical DSP methods to build our synthesizer, which allow for quick offline, and parallel generation of audio signals without the usage of GPUs. We made extensive use of Pippi\footnote{https://github.com/luvsound/pippi} and SciPy~\cite{jones2001scipy} libraries. Our virtual synthesizer contains a set of one or more submodules. Each submodule is a self-contained noise making unit. Submodules have identical sets of parameters, but widely different outputs can be achieved depending on the values assigned. The set of parameters available to each submodule is highlighted in Table~\ref{table:submodule_params}. The sonic output of the virtual synthesizer is the normalized addition of the sonic output of its submodules. Our implementation of a synthesizer can have any number of submodules. The parameters that dictate the output signal of each submodule as well as the range of values each parameter can take are shown in table~\ref{table:submodule_params}. We call the number of submodules in each virtual synthesizer the \textit{stack size}. We call the sets of parameter values that characterize a synthesizer's submodules a \textit{program} (analogous to a preset for a VST).

@@

We are not looking for the perfect imitation of organic drums using a synthesizer. We seek to imitate drums using a synthesizer and even \textbf{prefer} for its generations to retain novel, unusual characteristics. The task assigned to the virtual ear is the rare acceptance of drum-like sounds and the inevitable rejection of most \enquote{noise} outputs from the virtual synthesizer. The virtual ear makes two critical decisions: 
\emph{
\begin{quote}
\text{Decision.1 Could the sound be used as a drum?}\label{Decision.1}
\\
\text{Decision.2 If it does sound like a drum, what type of drum should it be?}\label{Decision.2}
\end{quote}
 }

\begin{figure}[htbp]
    \begin{center}
    \textbf{Data Overview}
    \makebox[\textwidth]{
    \includegraphics[width=0.95\linewidth]{images/venn_data.pdf}}
    \end{center}
    \caption{ 
    Scales are symbolic (true scales are not measurable and/or subjective).  $\mathcal{N}$ is the set of percussive sounds a synthesizer is capable of making. Our positive samples, $\mathcal{T^{+}}$, is a small fraction of a wide variety of percussive sounds that are conceivable. The size of $\mathcal{T^{-}}$ can be set to any number(we can generate as any number of random samples), but we prefer to keep it no larger than  $\mathcal{T^{+}}$. $\mathcal{H}$ is a series of sounds sent to the ear for classification. The ideal ear will detect  $\mathcal{H^{+}}$, the overlap of $\mathcal{H}$ and $\mathcal{N}$, with high precision and recall. 
    }
\label{fig:ven_data}
\end{figure}
\hypref{Decision.1} requires knowledge of what drums \textbf{do not} sound like, or knowledge of an infinitely large set, which cannot be fully represented via examples. An important consideration is that the source of sounds used in training the model (organic drum sounds) will be fundamentally different from the source of unlabeled sounds we wish to categorize (noise from a synthesizer). This issue is reflective of the OSR problem~\cite{geng2020recent,mundt2019open}. Figure~\ref{fig:ven_data} highlights a number of caveats with this approach. we denote $\mathcal{N}$ as the set of percussive sounds a synthesizer is capable of making. $\mathcal{T^{+}}$, our positive examples, is a set of sound extracted from material drums. $\mathcal{T^{-}}$, is a small subset of an infinite set which is meant to represent all noises which the synthesizer is capable of making and likely includes sounds in $\mathcal{N}$. During the hearing test, the virtual ear continually receives synthesizer noises for classification. The set of received noises is denoted as $\mathcal{H}$. The performance of the ear is measured by its precision and recall in finding the overlap between $\mathcal{H}$ and $\mathcal{N}$. It is important to ensure that the change in learning domains---particularly with the case of positive examples of percussive sounds---does not interfere with transformation of knowledge from the training of the ear to the hearing test. To maximize the transference of knowledge from one domain to another, we seek to learn from agnostic feature sets that capture fundamental characteristics of the data points.
\subsection{Feature Extraction}
To train our virtual ear, we require generalizable, domain agnostic features. Various works have demonstrated effective reconstruction of signals given their Short-time Fourier Transforms (STFT) ~\cite{nawab1983signal,griffin1984signal}. If the STFT of a signal can be used its reconstruction, perhaps it can be utilized as a source of fundamental features necessary for audible signal categorization. We defined 3 STFT transformation functions which we believe to capture important, unique attributes of percussive sounds (figure ~\ref{fig:stackspectrums}). Our transformations are GPU based, allowing their application in tandem to the training of the classifier or tuning of STFT transformation during hyper-paramter optimization (see Appendix~\ref{appendix:hyperparam}).
\subsection{Ear Types}
\label{classifiers}

@@

\subsection{Survey Discussion}
\label{survey2_takeaway}
Due to the discrepancy in categorization and output groups, results from the two surveys must be compared with caution. Agree-ability scores for MEM pipeline are often lower, but we reiterate that the MEM survey has two out of six (rather than one out of six in the TPE pipeline) categories which are exclusive to survey responders. Supporting this hypothesis is the notable improvement in agree-ability observed after dropping all \enquote{bad} and \enquote{other} samples. Despite the drawbacks, the MEM pipeline remains competitive while working with a fraction of the features to learn from and simultaneously making both decisions. Suggesting that latent representations of autoencoder networks can be used as general, low-dimensional proxies for representation of complex inputs.  
\bibliographystyle{apacite}
\bibliography{CSMC_MUME_LaTeX_Template.bib}
